\bf Calendar-based graphics for visualizing people's daily schedules

### 2018-10-22 main.Rmd
spell check and fixed bib

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the inner-city area till the end of 2016 (Figure \ref{fig:ped-map}). The dataset can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.

@@

```{r time-series-plot, fig.height = 6, fig.cap = "Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne. Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number. A spike occurred at the State Library, caused by the annual White Night event on 20th of February. A relatively fixed pattern repeats from one week to another at Flagstaff Station. Birrarung Marr looks rather random and spiky, with chunks of missing records."}
```

@@

To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors. Figure \ref{fig:facet} presents the idea of faceting calendar plots. This allows comparing the overall structure between sensors, while emphasising individual sensor variation. In particular, we can immediately learn that when Birrarung Marr was crowded and why, such as Australian Open in the last two weeks of January. This does not pop out in the conventional graphics.

@@

```{r h1, fig.height = 5, fig.cap = "Calendar display for household 1, indicates higher weekend usage, and in the summer months, November-February. It seems that they took a vacation in June."}
```

```{r h2, fig.height = 5, fig.cap = "Calendar display for household 2, reveals their tendency to use air conditioning and heating continuously. Not many vacation were taken."}
```

```{r h3, fig.height = 5, fig.cap = "Calendar display for household 3. Their energy use reveals higher energy use in the winter months, with multiple peaks daily on both week days and weekends. There are some high peaks in summer, perhaps indicating occasional air conditioner use. There have been several long vacations in the past year."}
```

```{r h4, fig.height = 5, fig.cap = "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, and June."}
```

@@

The methodology creates the western calendar layout, because most countries have adopted this format. The main difference between countries is the use of different languages for labelling, which is supported by the software. Layouts beyond the western calendar could be achieved by the same modular arithmetic approach.

>>>>

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the inner-city area till the end of 2016 (Figure \ref{fig:ped-map}). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.

@@

```{r time-series-plot, fig.height = 6, fig.cap = "Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne. Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number. A spike occurred at the State Library, caused by the annual White Night event on 20th of February. A relatively persistent pattern repeats from one week to another at Flagstaff Station. Birrarung Marr looks rather noisy and spiky, with chunks of missing records."}
```

@@

To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors. Figure \ref{fig:facet} presents the idea of faceting calendar plots. This allows comparing the overall structure between sensors, while emphasizing individual sensor variation. In particular, we can immediately learn that when Birrarung Marr was crowded and why, such as Australian Open in the last two weeks of January. This does not pop out in the conventional graphics.

@@

```{r h1, fig.height = 7, fig.cap = "Calendar display for household 1, indicates higher weekend usage, and in the summer months, November-February. It seems that they took a vacation in June."}
```

```{r h2, fig.height = 7, fig.cap = "Calendar display for household 2, reveals their tendency to use air conditioning and heating continuously. Not many vacation were taken."}
```

```{r h3, fig.height = 7, fig.cap = "Calendar display for household 3. Their energy use reveals higher energy use in the winter months, with multiple peaks daily on both week days and weekends. There are some high peaks in summer, perhaps indicating occasional air conditioner use. There have been several long vacations in the past year."}
```

```{r h4, fig.height = 7, fig.cap = "Calendar display for household 4, shows energy use mostly in the evenings and on weekends. Three short trips were taken in October, December, and June."}
```

@@

The methodology creates the western calendar layout, because most countries have adopted this format. The main difference between countries is the use of different languages for labeling, which is supported by the software. Layouts beyond the western calendar could be achieved by the same modular arithmetic approach.

### 2018-10-22 main.tex
spell check and fixed bib

The purpose of the calendar-based visualization is to provide insights
into people's daily schedules relative to events such as work days,
weekends, holidays, and special events. This work was originally
motivated by studying foot traffic in the city of Melbourne, Australia
\citep{ped}. There have been 43 sensors installed that count pedestrians
every hour across the inner-city area till the end of 2016 (Figure
\ref{fig:ped-map}). The dataset can shed light on people's daily
rhythms, and assist the city administration and local businesses with
event planning and operational management. A routine examination of the
data would involve constructing conventional time series plots to catch
a glimpse of temporal patterns. The faceted plots in Figure
\ref{fig:time-series-plot}, give an overall picture of the foot traffic
at three different sensors over 2016. Further faceting by day of the
week (Figure \ref{fig:facet-time}) provides a better glimpse of the
daily and sub-daily pedestrian patterns.

@@

\caption{Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne. Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number. A spike occurred at the State Library, caused by the annual White Night event on 20th of February. A relatively fixed pattern repeats from one week to another at Flagstaff Station. Birrarung Marr looks rather random and spiky, with chunks of missing records.}\label{fig:time-series-plot}
\end{figure}

>>>>

The purpose of the calendar-based visualization is to provide insights
into people's daily schedules relative to events such as work days,
weekends, holidays, and special events. This work was originally
motivated by studying foot traffic in the city of Melbourne, Australia
\citep{ped}. There have been 43 sensors installed that count pedestrians
every hour across the inner-city area till the end of 2016 (Figure
\ref{fig:ped-map}). The data set can shed light on people's daily
rhythms, and assist the city administration and local businesses with
event planning and operational management. A routine examination of the
data would involve constructing conventional time series plots to catch
a glimpse of temporal patterns. The faceted plots in Figure
\ref{fig:time-series-plot}, give an overall picture of the foot traffic
at three different sensors over 2016. Further faceting by day of the
week (Figure \ref{fig:facet-time}) provides a better glimpse of the
daily and sub-daily pedestrian patterns.

@@

\caption{Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne. Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number. A spike occurred at the State Library, caused by the annual White Night event on 20th of February. A relatively persistent pattern repeats from one week to another at Flagstaff Station. Birrarung Marr looks rather noisy and spiky, with chunks of missing records.}\label{fig:time-series-plot}
\end{figure}
